---
title: "yuzhunna"
author: "Yusuf Uzhunnan"
date: "2/15/2021"
output: html_document
---

```{r}
library(ggplot2)
library(dplyr)
```

```{r}
hotelbookings <- read.csv("hotel_bookings.csv")
summary(hotelbookings)
```


```{r}
# New col for total guests
hotelbookings$total_guests <- hotelbookings$adults + hotelbookings$children + hotelbookings$babies
hotelbookings$total_guestsfactor <- as.factor(hotelbookings$total_guests)

aggregate(is_canceled ~ adults, data=hotelbookings, mean)
aggregate(is_canceled ~ children, data=hotelbookings, mean)
aggregate(is_canceled ~ babies, data=hotelbookings, mean)

nrow(hotelbookings)
nrow(hotelbookings[hotelbookings$is_canceled == 1, ])

summary(hotelbookings$total_guests)
cancelratebyguests <- aggregate(is_canceled ~ total_guests, data=hotelbookings, mean)

cancelratebyguests

barplot(cancelratebyguests$is_canceled, col=cancelratebyguests$total_guests, border =TRUE, axes= TRUE, ylim=c(0,1))


barplot(tapply(hotelbookings$is_canceled,hotelbookings$total_guestsfactor, mean, na.rm = TRUE), border = TRUE, axes = TRUE, ylim = c(0,1))

barplot(cancelratebyguests$is_canceled, col=cancelratebyguests$total_guests, names.arg = cancelratebyguests$total_guests, xlab="Total Number of Guests", ylab="Cancellation Rate", border =TRUE, axes= TRUE, ylim=c(0,1))


```



```{r}
# Distribution Channel

canceledbydistchannel <- tapply(hotelbookings$is_canceled, hotelbookings$distribution_channel, mean, na.rm = TRUE)
barplot(sort(canceledbydistchannel, decreasing = TRUE),ylab = "Cancellation Rate",xlab = "Distribution Channel", ylim = c(0,1), col = c("red","blue","green", "pink","yellow"))

canceled <- hotelbookings[hotelbookings$is_canceled == 1, ]
p<-ggplot(canceled, aes(x=distribution_channel, y=is_canceled)) +
  geom_bar(stat="identity")+theme_minimal()
p
```

From the graph above, we can see that the "Undefined" distribution channel has the highest cancellation rate.However, when we look at quantity canceled for the Undefined channel in the second graph, we see that there are very few total "Undefined" cancellations despite the rate being very high. Aside from this, the channel with the highest cancellation rate is Travel Agents and Tour Operations.This distribution channel also has the highest total number of cancellations.



```{r}
# relationship btwn # of special requests and total stay length, color coded by is_canceled
hotelbookings$totalstaylength <- hotelbookings$stays_in_week_nights + hotelbookings$stays_in_weekend_nights
ggplot(data=hotelbookings, aes(x=totalstaylength, y=total_of_special_requests, colour=is_canceled)) + geom_point(size=2, alpha=.5) + geom_smooth(method = "lm") + geom_jitter()
```

From the graph above, there are two conclusions that can be made. First, shorter stays seem to be linked with more special requests. Second, longer stays seem to have more cancellations than shorter stays. Based on this information, shorter stays should be targeted as special requests are higher (which could bring additional revenue) and there also tends to be less cancellations.