---
title: "Data Visualization"
author: "Anthony Zhao"
date: "4/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import Data

```{r}
stroke <- read.csv("healthcare-dataset-stroke-data.csv")
stroke <- stroke[stroke$gender != "Other",]
stroke$id <- NULL
stroke$gender <- as.factor(stroke$gender)
stroke$work_type <- as.factor(stroke$work_type)
stroke$smoking_status <- as.factor(stroke$smoking_status)
stroke$ever_married <- as.factor(stroke$ever_married)
stroke$Residence_type <- as.factor(stroke$Residence_type)
stroke$bmi <- as.numeric(stroke$bmi)
summary(stroke)
```

## Process Data

```{r}
# create the bmi data that doesn't have NA
bmi_data <- stroke[!is.na(stroke$bmi),]
# can't use stroke to predict bmi
bmi_model <- lm(bmi ~ .-stroke, data = bmi_data)
summary(bmi_model)
# remove insignificant variables
bmi_model <- lm(bmi ~ .-stroke-gender-Residence_type+age*hypertension*ever_married-heart_disease, data = bmi_data)
summary(bmi_model)

stroke$bmi[is.na(stroke$bmi)] = predict(bmi_model, stroke[is.na(stroke$bmi),])

normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

strokemm <- as.data.frame(model.matrix(~.-1,stroke))
stroke_norm <- as.data.frame(lapply(strokemm, normalize))
summary(stroke_norm)
```

## Visualizations
```{r}
par(mfrow=c(2,2))
barplot(prop.table(table(stroke$stroke, stroke$gender), 2), main="Gender vs Stroke", xlab="Gender", col=c("darkblue","red"))
barplot(prop.table(table(stroke$stroke, stroke$hypertension), 2), main="Hypertension vs Stroke", xlab="Hypertension", col=c("darkblue","red"))
barplot(prop.table(table(stroke$stroke, stroke$heart_disease), 2), main="Heart Disease vs Stroke", xlab="Heart Disease", col=c("darkblue","red"))
barplot(prop.table(table(stroke$stroke, stroke$ever_married), 2), main="Marriage vs Stroke", xlab="Marriage", col=c("darkblue","red"))
par(mfrow=c(2,2))
barplot(prop.table(table(stroke$stroke, stroke$work_type), 2), main="Work vs Stroke", xlab="Work", col=c("darkblue","red"))
barplot(prop.table(table(stroke$stroke, stroke$Residence_type), 2), main="Residence vs Stroke", xlab="Residence Type", col=c("darkblue","red"))
barplot(prop.table(table(stroke$stroke, stroke$smoking_status), 2), main="Smoker vs Stroke", xlab="Smoking Status", col=c("darkblue","red"))
library(lsr)
barplot(prop.table(table(stroke$stroke, quantileCut(stroke$age, 10)), 2), main="Age vs Stroke", xlab="Age", col=c("darkblue","red"))
par(mfrow=c(2,2))
barplot(prop.table(table(stroke$stroke, quantileCut(stroke$bmi, 10)), 2), main="BMI vs Stroke", xlab="BMI", col=c("darkblue","red"))
barplot(prop.table(table(stroke$stroke, quantileCut(stroke$avg_glucose_level, 10)), 2), main="Glucose vs Stroke", xlab="Glucose Level", col=c("darkblue","red"))
```

## Inter-Variable
```{r}
prop.table(table(stroke$ever_married, stroke$hypertension), 1)
prop.table(table(stroke$ever_married, stroke$heart_disease), 1)
prop.table(table(stroke$hypertension, stroke$heart_disease), 1)
par(mfrow=c(2,2))
barplot(prop.table(table(stroke$ever_married, quantileCut(stroke$age, 10)), 2), main="Age vs Marriage", xlab="Age", col=c("darkblue","red"))
barplot(prop.table(table(stroke$hypertension, quantileCut(stroke$age, 10)), 2), main="Age vs Hypertension", xlab="Age", col=c("darkblue","red"))
barplot(prop.table(table(stroke$heart_disease, quantileCut(stroke$age, 10)), 2), main="Age vs Heart Disease", xlab="Age", col=c("darkblue","red"))
par(mfrow=c(2,2))
barplot(prop.table(table(stroke$heart_disease[stroke$ever_married=="Yes"], cut(stroke$age[stroke$ever_married=="Yes"], breaks = c(0,10,20,30,40,50,60,70,80,90))), 2), main="Age vs Heart Disease for Married", xlab="Age", col=c("darkblue","red"))
barplot(prop.table(table(stroke$heart_disease[stroke$ever_married=="No"], cut(stroke$age[stroke$ever_married=="No"], breaks = c(0,10,20,30,40,50,60,70,80,90))), 2), main="Age vs Heart Disease for Unmarried", xlab="Age", col=c("darkblue","red"))
barplot(prop.table(table(stroke$hypertension[stroke$ever_married=="Yes"], cut(stroke$age[stroke$ever_married=="Yes"], breaks = c(0,10,20,30,40,50,60,70,80,90))), 2), main="Age vs Hypertension for Married", xlab="Age", col=c("darkblue","red"))
barplot(prop.table(table(stroke$hypertension[stroke$ever_married=="No"], cut(stroke$age[stroke$ever_married=="No"], breaks = c(0,10,20,30,40,50,60,70,80,90))), 2), main="Age vs Hypertension for Unmarried", xlab="Age", col=c("darkblue","red"))

```

Although it seemed liked marriage, hypertension, and heart disease all lead to higher chances of stroke, age clearly has a very large impact on these three variables.

When we split up based on marriage and age, we can see people of each age group are not much more prone to heart disease or hypertension due to marriage.